---
title: "PA1_template"
author: "erikaananda"
date: "June 13, 2015"
output: html_document 
---
## The following is an analysis of fitness metrics captured by activity monitoring devices 

Read the activity.csv file

```{r}
library(lubridate)
library(dplyr)
library(lattice)
#setwd("./Documents/RData/data/")
DSActivity <- read.csv("activity.csv")
```
 
Process the Data Frame to recognize dates and remove NAs.

```{r}
DSActivity <- mutate (DSActivity, date = ymd(DSActivity$date))
CompleteActivity <- DSActivity[complete.cases(DSActivity), ]
```
#Totals By Day
Get a summary of steps per date and plot a histogram
```{r}
par(mfrow=c(1,1))
GDateActivity <- group_by(CompleteActivity, date)
StepsPerDate <- 
  summarize(GDateActivity, sum(steps), mean(steps), median(steps))
names(StepsPerDate) = 
  c("date", "totalSteps", "meanSteps", "medianSteps")
hist(x= StepsPerDate$totalSteps, main = "Histogram of Total Steps", xlab = "Total Steps")
```

Take a look at the mean and median of the total number of steps taken per day
```{r}

plot(StepsPerDate$date, StepsPerDate$meanSteps, 
     col = "red", type = "l", main = "Average Steps Per Day", 
     xlab = "Day", ylab= "Average Steps")

```

Our median steps per day comes to 0
```{r}
unique(StepsPerDate$medianSteps)
```
#Daily Patterns
Plot the Average Steps per Interval
```{r}
GIntActivity <- group_by(CompleteActivity, interval)
StepsPerInterval <- 
  summarize(GIntActivity, sum(steps), mean(steps), median(steps))

names(StepsPerInterval) = 
  c("interval", "totalSteps", "meanSteps", "medianSteps")
  
plot(as.integer(StepsPerInterval$interval), StepsPerInterval$meanSteps, 
     col = "red", type = "l", main = "Average Steps Per Interval", 
     xlab = "Interval", ylab= "Average Steps")
```

Which interval has the most steps?
```{r}
max(StepsPerInterval$totalSteps)
```
#Missing Values
How many missing values are there?
```{r}
sum(!complete.cases(DSActivity))
```
Fill missing values in with the mean for that Interval
```{r}
ImputeActivity <- DSActivity

for (i in 1:nrow(ImputeActivity)) {
  if (is.na(ImputeActivity[i, 1])) {
    intervalMean <- StepsPerInterval[StepsPerInterval$interval == ImputeActivity[i, 3], 3]
    ImputeActivity[i, 1] <- intervalMean
    }
}

```
Create a new histogram with imputed data
```{r}
IGDateActivity <- group_by(ImputeActivity, date)
IStepsPerDate <- 
  summarize(IGDateActivity, sum(steps), mean(steps), median(steps))
names(IStepsPerDate) = 
  c("date", "totalSteps", "meanSteps", "medianSteps")
hist(x= IStepsPerDate$totalSteps, main = "Histogram of Total Steps with Imputed Values", xlab = "Total Steps")
```
Look above to compare with the data where NAs are removed - 
the Y axis is much shorter

Now, for the mean and median and compare with the original. 
The median is seriously affected.
```{r}
par(mfrow=c(2,2))
plot(as.integer(StepsPerDate$date), StepsPerDate$meanSteps, 
     col = "red", type = "l", main = "Average Steps Per Day", 
     xlab = "Day", ylab= "Average Steps")

plot(as.integer(StepsPerDate$date), StepsPerDate$medianSteps, 
     col = "red", type = "l", main = "Median Steps Per Date", 
     xlab = "Day", ylab= "Average Steps")
plot(as.integer(IStepsPerDate$date), IStepsPerDate$meanSteps, 
     col = "red", type = "l", main = "Imputed Average Steps Per Date", 
     xlab = "Day", ylab= "Average Steps")

plot(as.integer(IStepsPerDate$date), IStepsPerDate$medianSteps, 
     col = "red", type = "l", main = "Imputed Median Steps Per Date", 
     xlab = "Day", ylab= "Average Steps")
```

#Weekdays Vs Weekends
```{r}
IWeekday <- weekdays(ImputeActivity$date)
for (i in 1:length(IWeekday)){
  if (IWeekday[i] == "Saturday" | IWeekday[i] ==  "Sunday"){
    IWeekday[i] <- "weekend"
  }
  else IWeekday[i] <- "weekday"
}
ImputeActivity <- mutate(ImputeActivity, dayType = IWeekday)

IGIntActivity <- group_by(ImputeActivity, interval, dayType)
StepsPerInterval <- 
  summarize(IGIntActivity, sum(steps), mean(steps), median(steps))


names(StepsPerInterval) = 
  c("interval", "dayType", "totalSteps", "meanSteps", "medianSteps")
  

xyplot(meanSteps ~ interval | dayType, data = StepsPerInterval,
       layout = c(1, 2), ylab = "Average Number of Steps", 
       type = "l") ## Plot with 2 panels



```
